%% ─────────────────────────────────────────────
\section{LLM Command Layer}
The LLM command layer endows the BDI coordinator with natural-language directive processing under a strict division of labour: the model decides \textit{what} to do; the BDI plan library decides \textit{how}. The coordinator is the only agent with an LLM instance; the worker executes structured JSON orders over the chat channel.

\subsection{Routing, Classification, and Directive Lifecycle}
Every incoming message is routed through a short-circuit chain before a model call is made. Partner JSON (starts with \texttt{\{}) is consumed by the protocol handler without classification. Abort keywords skip the queue and call \texttt{abortCurrent()} immediately. Messages ending with \texttt{?} or matching greeting patterns are forwarded directly to \texttt{handleChat()} — no classify call. Only unresolved messages reach \texttt{classifyDirective}, which returns one of five labels: \texttt{STOP} (freeze both agents), \texttt{GO} (resume), \texttt{ACTION} (serialized lane), \texttt{CHAT} (concurrent read-only lane), or \texttt{IGNORE} (a bare ``red light''/``green light'' with no stop/resume imperative and not an announcement --- dropped in \texttt{route()} before any handling, so it never arms, stops, or resumes the agents, even mid-game). The classifier replaced an earlier anchored-regex approach because the same words appear in opposite contexts: the live ``RED LIGHT!'' command (\texttt{STOP}) and the mission announcement ``\textit{Let's begin a red light green light game}'' (\texttt{ACTION}, routing the LLM to call \texttt{start\_light\_mission}). \texttt{STOP}/\texttt{GO} signals only take effect after \texttt{lightMission.active} is armed, preventing stray mentions from freezing the agents. A \texttt{pending} slot (not a queue) ensures the latest directive always wins.

\texttt{runDirective} runs a flat ReAct loop (up to 30 iterations). A pre-computed planner was rejected: BDI tools abstract navigation into atomic calls (\texttt{deliver()} navigates and scores), so plan depth rarely exceeds 3--5 steps, and pre-computed plans expire the instant a parcel is claimed by a competitor. Action directives terminate via an \texttt{End} marker on the model's last action — not a \texttt{Final Answer} — preventing post-action confirmations from reaching chat. When a tool in the \texttt{MISSION\_TOOLS} set succeeds, the loop auto-emits ``Mission accepted.'' on directive end; when a Level-3 routine refuses a net-penalty offer, it emits ``Mission declined.'' instead. Both resolve a live bug where missions were applied or refused silently. \texttt{MAX\_TOOL\_FAILURES}~= 1 triggers abort and BDI handback after a single failed command.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    >={Stealth[round]},
    every node/.style={font=\scriptsize, align=center},
    box/.style={draw, rounded corners, fill=black!4, text width=28mm, minimum height=10mm},
    init/.style={draw, fill=blue!5, text width=28mm, minimum height=10mm},
    proc/.style={draw, fill=white, text width=28mm, minimum height=10mm},
    gate/.style={draw, diamond, aspect=1.6, inner sep=2pt, fill=white},
    flow/.style={->, thick},
    off/.style={->, thick, densely dashed, gray!80}
]

%% ═══════════════════════════ NODE PLACEMENT ═══════════════════════════

%% ── Column 0: Intake ──
\node[box]  at (0.0,  0)    (msg)  {Incoming\\message};
\node[gate] at (2.8,  0)    (pre)  {route?};

%% ── Column 1: Tool invocation ──
\node[proc] at (2.8, -9.8)  (tool) {invoke tool()\\observation};

%% ── Column 2: Classify + ReAct loop ──
\node[init] at (5.2,  0)    (cls)  {\texttt{classifyDirective}\\1 LLM call};
\node[gate] at (8.0,  0)    (lbl)  {label?};
\node[box]  at (8.0, -2.2)  (pen)  {\texttt{pending} slot\\latest-wins,\\abort if busy};
\node[init] at (8.0, -4.2)  (rd)   {\texttt{runDirective}\\flat ReAct\\MAX\_ITER = 30};
\node[proc] at (8.0, -6.2)  (llm)  {\texttt{callModel()}};
\node[gate] at (8.0, -8.2)  (prs)  {response?};
\node[box]  at (8.0,-10.5)  (endn) {End marker\\silent / Mission\\accepted./declined.};

%% ── Column 3: Fast-lane + autonomy gate ──
\node[box]  at (12.2,  1.0) (sg)   {halt or resume\\both agents};
\node[proc] at (12.2, -0.8) (chat) {\texttt{runConversation}\\read-only,\\fast-lane};
\node[box]  at (12.2, -4.2) (act)  {\texttt{directive.active}\\true on cmd entry\\false in finally};
\node[init] at (12.2, -6.2) (bdi)  {BDI plan library\\A\textsuperscript{*} / PDDL};
\node[proc] at (12.2, -9.8) (fin)  {Final Answer\\reply to sender};


%% ═══════════════════════════ EDGE ROUTING ═══════════════════════════

%% ── Intake ──
\draw[flow] (msg) -- (pre);

%% ── Pre-route: three exits from three anchors ──
%% west  → partner JSON / abort kw (dead-end, out of frame)
\draw[off]  (pre.west) -- ++(-1.6, 2.2)
            node[above left, font=\tiny]{\shortstack{partner JSON\\abort kw}};
%% north → ?/greeting: arc over the top directly to runConversation
\draw[flow] (pre.north) -- ++(0,5.8) -|
            node[pos=0.06, right]{\tiny ?/greeting} (chat.north);
%% east  → classifyDirective (normal path)
\draw[flow] (pre.east) -- node[above]{\tiny else} (cls);

%% ── Classify → label ──
\draw[flow] (cls) -- (lbl);

%% ── Label branches ──
\draw[flow] (lbl.north) |- node[pos=0.5, above]{\tiny STOP/GO}             (sg.west);
\draw[flow] (lbl.east)  |- node[pos=0.5, right, yshift=4pt]{\tiny CHAT}    (chat.west);
\draw[flow] (lbl.south) -- node[right]{\tiny ACTION}                        (pen.north);

%% ── ReAct core ──
\draw[flow] (pen) -- (rd);
\draw[flow] (rd)  -- (llm);
\draw[flow] (llm) -- (prs);

%% ── Response branches ──
\draw[flow] (prs.west)  -| node[pos=0.2, above]{\tiny action} (tool.north);
\draw[flow] (prs.south) --  node[right]{\tiny end}             (endn.north);
\draw[flow] (prs.east)  -| node[pos=0.2, above]{\tiny final}  (fin.north);

%% ── Observation feedback: tool → llm ──
\draw[off]  (tool.north) -- ++(0,3.1) -|
            node[pos=0.07, right]{\tiny observation} (llm.west);

%% ── Autonomy gate ──
\draw[flow] (rd.east)  -- node[above]{\tiny cmd tool} (act.west);
\draw[flow] (act)      --                              (bdi);
\draw[off]  (bdi.east) -- ++(0.5,0) |- node[pos=0.25,right]{\tiny done} (act.east);

%% ── Next directive outer loop ──
\draw[off] (endn.south) -- ++(0,-0.6) -|
           node[pos=0.1, below]{\tiny next directive} (15.0,-2.2) -- (pen.east);

\end{tikzpicture}
\caption{LLM layer execution flow: pre-route short-circuits, ReAct loop, and autonomy gate.}
\label{fig:llmloop_clean}
\end{figure}

\subsection{Autonomy Gate: Optimistic Release}
Initially the gate (\texttt{directive.active}) was held across the full directive, suppressing BDI through every inter-command model call. The coordinator would idle at a delivery tile for a full LLM round-trip instead of resuming parcel work. The gate is now released in each command's \texttt{finally} block:
\begin{verbatim}
directive.active = true;       // re-take on entry
commandAndAwait(predicate);    // BDI executes the intention
finally { directive.active = false; resumeAutonomy?.(); }
\end{verbatim}
A follow-up command re-acquires the gate and calls \texttt{haltCurrent()} to cancel any BDI work started in the gap. Two stationary commands manage their own gate without going through this helper: \texttt{wait(N)} holds \texttt{directive.active} for the sleep duration and releases it on exit; \texttt{hold()} sets \texttt{manualHold.active = true} and persists \textit{after} the directive ends (it survives the directive's \texttt{finally}), keeping the agent frozen until \texttt{release\_hold()} is called. \texttt{handoffRunning()} suppresses the gate release in the directive's \texttt{finally} so the background handoff loop retains the gate beyond the directive that started it.

\subsection{Tool Design: BDI–LLM Interface}
Tools are split into categories by their relationship to the autonomy gate (Table~\ref{tab:tools}). \textbf{Reasoning} and \textbf{Read} tools have no world effect and run safely on the concurrent CHAT lane. \textbf{Command} tools push a BDI intention via \texttt{commandAndAwait} and block until the plan library completes it — the LLM supplies the intent, A*/PDDL supplies the execution path. \textbf{Partner} tools relay orders to the worker without gating this agent. \textbf{Mission} tools apply or remove constraints; on success the loop emits ``Mission accepted.''

\begin{table}[H]
\centering\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Category} & \textbf{Tools} & \textbf{Gate effect} \\
\midrule
Reasoning  & \texttt{calculate}, \texttt{get\_current\_time} & none \\
Read       & \texttt{get\_my\_position}, \texttt{sense\_parcels}, \texttt{sense\_delivery\_tiles}, & none \\
           & \texttt{sense\_spawn\_tiles}, \texttt{get\_map\_info}, \texttt{path\_cost}, \texttt{ask\_partner\_status} & \\
Command    & \texttt{go\_to}, \texttt{go\_to\_stay}, \texttt{go\_pickup}, \texttt{deliver}, & take/release gate \\
           & \texttt{put\_down}, \texttt{go\_put\_down} & \\
           & \texttt{pickup\_next\_parcel} & release, then re-take \\
           & \texttt{wait} & hold for duration \\
           & \texttt{hold} / \texttt{release\_hold} & persist past directive \\
           & \texttt{say} & none (chat only) \\
Partner    & \texttt{order\_partner\_goto/pickup/deliver/putdown} & none (worker moves) \\
           & \texttt{halt\_partner}, \texttt{resume\_partner} & none \\
Mission L2 & \texttt{apply\_mission}, \texttt{forbid\_delivery}, \texttt{restrict\_exploration} & none \\
           & \texttt{dropMission}, \texttt{dropMissions} & none \\
Mission L3 & \texttt{start\_handoff} / \texttt{stop\_handoff} & handoff owns gate \\
           & \texttt{gather\_near} & take/release gate \\
           & \texttt{start\_light\_mission} / \texttt{stop\_light\_mission} & none \\
\bottomrule
\end{tabular}
\caption{LLM tool catalogue grouped by gate effect.}
\label{tab:tools}
\end{table}

Three tools embody distinct points on the BDI–LLM control spectrum:
\begin{itemize}
    \item \textbf{\texttt{pickup\_next\_parcel}}: releases \texttt{directive.active} and polls until a new parcel ID appears in \texttt{parcels.carriedBy(me.id)}, then re-takes the gate and halts any in-progress BDI intention. The BDI strategy handles trap avoidance, decay scoring, and spawner touring — capabilities no LLM polling loop can match without saturating the proxy.
    \item \textbf{\texttt{gather\_near(x,y,D)}}: encodes all geometry in code — reachability BFS per agent, selection of two distinct walkable tiles within distance~$D$, worker parking before B moves, coordinator movement, \texttt{manualHold} on both — so the model issues one call instead of a fragile multi-step sequence prone to picking walls or occupied tiles. A pre-call \texttt{requestStatus()} refreshes the worker's live position, since it only streams while under an order.
    \item \textbf{\texttt{forbid\_delivery}}: resolves a side keyword (\texttt{leftmost}/\texttt{rightmost}/\texttt{top}/\texttt{bottom}) or explicit coordinates to actual delivery tiles, then narrows \texttt{allowedDeliveryTiles} additively. It refuses if the result would leave zero deliverable tiles, preventing the agent from being stranded.
\end{itemize}
\texttt{path\_cost(x,y)} returns a JSON object \texttt{\{steps, estSeconds, decayLostPerCarriedParcel\}} computed from A* path length and EMA-tracked real move timing. Live position and cargo count are injected into every observation so the model cannot act on stale carry state (e.g., issuing \texttt{deliver()} after a \texttt{put\_down()} that already emptied its hands). The navigation tools \texttt{go\_to}, \texttt{go\_to\_stay}, and \texttt{go\_put\_down} additionally accept a side keyword (\texttt{leftmost}/\texttt{rightmost}/\texttt{top}/\texttt{bottom}), resolved over the reachable tiles --- like \texttt{get\_map\_info} --- to the one \textit{nearest} the agent; \texttt{go\_put\_down} is the navigate-and-drop sibling of \texttt{go\_pickup}/\texttt{deliver}, collapsing the former \texttt{get\_map\_info}~$\to$~\texttt{go\_to\_stay}~$\to$~\texttt{put\_down} chain into one call.

\subsection{Mission Evaluation: Wiring Points into the BDI Cost Function}
Deciding whether a mission is \textit{worth doing} is the crux of challenge 2: special missions score more than ordinary delivery but may not pay for the work they cost. Rather than have the LLM judge worth from English, the system pushes a mission's literal $+N$/$-N$ points onto the \textbf{same value scale the BDI cost function already maximizes}. That cost function (\texttt{Strategy.js}) ranks options in points net of travel decay --- \texttt{bankNowValue()}~$=\texttt{scale}\cdot R - n\rho d_0$ (deliver the current load now), \texttt{pickupValue()} (detour to grab one more), \texttt{bankFirstValue()} --- so a mission's reward must compete head-to-head with parcel income.

The model's first-pass screen uses \textbf{\texttt{path\_cost(x,y)}} to compare the stated bonus against \texttt{decayLostPerCarriedParcel}~$\times$~\texttt{steps} (approximate forgone income per detour step): a 700-pt bonus 8 steps away is an easy accept, a 5-pt one a decline. But the binding decision is numeric and made by the BDI, not the prompt. A \texttt{oneShotBonus} go-there reward is valued by \texttt{bonusGoalValue()}~$= \texttt{points} - n\rho d$ and diverted to only when \texttt{bonusDiversion()} finds it beats banking the current load by a margin (\texttt{net}~$-$~\texttt{bankNowValue()}~$>$~\texttt{SWITCH\_MARGIN}); this hook runs once before each \texttt{decide()}, so every strategy is bonus-aware and a negative ``bonus'' is simply never chosen --- the penalty is avoided as a \textit{byproduct of maximization}, not a hard rule.

Accepted missions map to two levels. \textbf{Level-2} persistent constraints (\texttt{apply\_mission}, \texttt{forbid\_delivery}, \texttt{restrict\_exploration}) are enforced by \texttt{Strategy.js} gates wired into all strategy variants so they survive autonomy-gate transitions. Stacking is a FLOOR/CAP pair --- \texttt{requiredStackSize} (``at least $N$''), \texttt{maxStackSize} (``at most $N$''), both set for ``exactly $N$''; \texttt{deliveryMultipliers} scale a tile's delivery reward inside \texttt{bankNowValue}/\texttt{pickupValue}; \texttt{penaltyTiles} record the magnitude and fold the key into \texttt{avoidTiles} so A* never routes there. Every constraint is mirrored to the worker via \texttt{applyAndMirror}, which pairs \texttt{applyMissionConfig} with \texttt{sendConstraint('apply', cfg)} — forgetting the mirror silently desyncs the two agents, a live bug fixed by the wrapper. \textbf{Level-3} routines (\texttt{start\_handoff}, \texttt{gather\_near}, \texttt{start\_light\_mission}) have no single tile to price, so each keeps a signed running net (\texttt{handoffNet}/\texttt{gatherNet}/\texttt{lightNet}): every fresh offer \textit{adds} its $\pm N$ to the type's total, and \texttt{armedByNet} arms the routine while net~$\geq 0$. Cumulative penalties thus stop a running routine and a later positive offer re-arms it without explicit cancel logic (e.g.\ a $-500$ then a $+1000$ offer of the same mission nets $+500$ and arms).

\subsection{Cross-Agent Handoff Routine}
\texttt{handoff.js} implements the ``one picks up, another delivers'' mission via an emergent meet-in-the-middle protocol. The coordinator (B) drives the map's chosen strategy to acquire parcels — same decay-weighted decisions as autonomous mode — until the strategy stops returning \texttt{go\_pick\_up} while loaded, or the next parcel's nearest-delivery distance exceeds the current one by \texttt{HANDOFF\_PASS\_MARGIN}~= 3 tiles (the pickup would carry the load past the drop). B then re-steers the worker (A) toward the live geometric midpoint of their positions every navigation pass, computed over the static walkable map so narrow corridors do not pin A onto B's delivery line. B drops only on Manhattan-adjacency to A detected from live sensing (\texttt{otherAgents}), not the streamed status which lags by one tile — trusting the stream alone made B miss the meet by one tile and detour around A as an obstacle (fixed in a live-test iteration). After the drop, A delivers detached while B restarts acquisition; a \texttt{deliveryInFlight} gate suppresses the drift-positioning orders during the active delivery to prevent a drift order from superseding A's in-flight pickup→deliver chain. If no adjacency is detected within \texttt{MEET\_TIMEOUT\_MS}~= 25\,s, B freezes A with \texttt{sendHalt}, keeps the load, and restarts the cycle. A deadlock-break parks the frozen worker on a delivery tile when it is the sole obstacle between B and the spawners.
%% ─────────────────────────────────────────────
